---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



1.  Loading the required libraries


```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(knitr)
library(png)
```

2.  Unzipping,reading, and processing data:

```{r}
unzip("activity.zip")
activity<-read.csv("activity.csv")
activity$date=as.Date(activity$date, '%Y-%m-%d')
```



3. Creating a dataframe of total number of steps taken each day:


```{r}
activity$date=as.Date(activity$date, '%Y-%m-%d')
by_day<-group_by(activity, date)
sum_steps<-summarize(by_day, steps=sum(steps), .groups = 'drop')
head(sum_steps)
```
4. Plotting a histogram for total number of steps taken each day


```{r}
ggplot(sum_steps,aes(x=date,y=steps)) +
        geom_bar(stat="identity", color="green") +
        scale_x_date( date_breaks = "3 days", date_labels = "%d-%b")+
        theme(axis.text.x=element_text(angle=90))
```

5. Calculating mean of total daily steps:

```{r}
mean(sum_steps$steps, na.rm=TRUE)
```

6. Calculating median of total daily steps:

```{r}
median(sum_steps$steps, na.rm=TRUE)
```

7. Creating a dataframe of average number of steps per each interval across all days:


```{r}

by_interval<-group_by(activity, interval)
avg_steps<-summarize(by_interval, steps=mean(steps,na.rm=TRUE), .groups = 'drop')
head(avg_steps)
```

8. Plotting average number of steps per each interval across all days:

```{r}
ggplot(avg_steps,aes(x=interval,y=steps)) +
        geom_line(color="blue") + 
        scale_x_continuous(breaks=seq(0,2400,100)) +
        scale_y_continuous(breaks=seq(0,225,25)) +
        theme(axis.text.x=element_text(angle=90))
```


9. Calculating the interval with the maximum number of steps (average across all days)


```{r}
avg_steps[avg_steps$steps==max(avg_steps$steps),]
```
10.  Calculating number of NA values for steps:
```{r}
nrow(activity)*mean(is.na(activity$steps))
```

11. Replacing NA values with mean values for the interval 


```{r}

activity2<-activity
for (row in 1:nrow(activity2)) {
        if (is.na(activity2[row,1])) {
                na_interval<-activity2[row,3]
                avg<-mean(activity2[activity2$interval==na_interval,1], na.rm=TRUE)
                activity2[row,1]<-avg
        }
}
```



12. Creating a dataframe of total number of steps taken each day (with replaced NA):


```{r}
activity2$date=as.Date(activity2$date, '%Y-%m-%d')
by_day2<-group_by(activity2, date)
sum_steps2<-summarize(by_day2, steps=sum(steps), .groups = 'drop')
head(sum_steps2)
```
13. Plotting a histogram for total number of steps taken each day (with NA replaced)


```{r}
ggplot(sum_steps2,aes(x=date,y=steps)) +
        geom_bar(stat="identity", color="green") +
        scale_x_date( date_breaks = "3 days", date_labels = "%d-%b")+
        theme(axis.text.x=element_text(angle=90))
```

14. Calculating mean of total daily steps (with replaced NA):

```{r}
mean(sum_steps2$steps, na.rm=TRUE)
```

15. Calculating median of total daily steps (with replaced NA):

```{r}
median(sum_steps2$steps, na.rm=TRUE)
```

16. Adding factor variable for weekend/weekdays values:

```{r}
activity2$day<-factor(if_else(weekdays(activity2[,2]) %in% c('Saturday','Sunday'),'weekend', 'weekday'))
str(activity2$day)
```

17. Creating a dataframe of average number of steps per each interval for weekdays:

```{r}

by_weekday_int<-group_by(activity2, interval,day)
avg_weekday<-summarize(by_weekday_int, steps=mean(steps), .groups = 'drop')
head(avg_weekday)
```

18. Comparing average number of steps per each interval for weekends and weekdays:


```{r}
ggplot(avg_weekday, aes(x=interval,y=steps, color=day)) +
        geom_line() + 
        scale_x_continuous(breaks=seq(0,2400,100)) +
        scale_y_continuous(breaks=seq(0,225,25)) +
        theme(axis.text.x=element_text(angle=90))+
  facet_wrap(~day,  ncol=1)
```

